```{r setup}
#| message: false
#| warning: false
library(sf)
library(terra)
library(ggplot2)
```

# Tree height {.unnumbered}

## Data 

This dataset was downloaded from Lang, N., Rodríguez, A. C., Schindler, K., & Wegner, J. D. (2021). Canopy top height and indicative high carbon stock maps for Indonesia, Malaysia, and Philippines (Version 1.0) [Data set]. Zenodo. http://doi.org/10.5281/zenodo.5012448

Tree height was estimated by combining Sentinel-2 optical satellite images and GEDI lidar data, using a deep convolutional neural network.

```{r}
dir_height <- "data/tree_height"
area <- list.files("data/umrbpl/", full.names = TRUE, pattern = "shp$") |>
  read_sf() |>
  st_transform(crs = "EPSG:4326")
if (!dir.exists(dir_height)) dir.create(dir_height)
if (!file.exists(paste0(dir_height, "/canopy_top_height_2020_umrbpl.tif"))) {
  zen4R::download_zenodo(
    "10.5281/zenodo.5012447", dir_height,
    "canopy_top_height_2020_philippines.tif"
  )
  paste0(dir_height, "/canopy_top_height_2020_philippines.tif") |>
    rast() |>
    crop(area) |>
    mask(area) |>
    writeRaster(paste0(dir_height, "/canopy_top_height_2020_umrbpl.tif"))
}
height <- rast(paste0(dir_height, "/canopy_top_height_2020_umrbpl.tif"))
```


## Comparison with TMF

```{r}
tmf <- rast("data/tmf/JRC_TMF_AnnualChange_v1_2020_ASI_ID76_N20_E120.tif") |>
  crop(area) |>
  mask(area) |>
  resample(height, method = "near")
types <- c(
  "Undisturbed", "Degraded", "Deforested",
  "Regrowth", "Water", "Other"
)
height_type <- lapply(1:6, function(i) {
  x <- ifel(tmf == i, 1, NA) |> mask(x = height)
  data.frame(type = types[i], value = x[!is.na(x)])
}) |>
  do.call(what = rbind)
```

```{r}
ggplot(height_type) +
  geom_boxplot(aes(x = type, y = canopy_top_height_2020_philippines)) +
  labs(x = NULL, y = "Canopy height (m)")
```

